Towards Automated Merging of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-05-03 sattose/paper.tex
Updates from Overleaf

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolpceurws}
\usepackage{float}
\usepackage{hyperref}

@@

\begin{abstract}
Duplication in source code can have a \todo{be more specific: what are examples of major impact?}major negative impact on the maintainability of source code. There are several techniques that can be used in order to merge clones, reduce duplication, improve the design of the code and potentially also reduce the total volume of a software system. In this study, we look into the opportunities to aid in the process of refactoring these duplication problems for object-oriented programming languages. We look into \todo{why?} redefinitions for different types of clones that have been used in code duplication research for many years. \todo{turn it around: link the need for refactoring aids to redefining} These redefinitions are aimed towards flagging only clones that are useful for refactoring purposes. \todo{do you already refer to our measurements?} Our current measurements indicate that more than half of the duplication in code is related to each other through inheritance, making it easier to refactor these clones in a clean way. Approximately a fifth of the duplication can be refactored through method extraction, while other clones require other techniques to be applied. More measurements will be conducted to get a detailed overview of in what contexts clones occur, and what this means for the refactoring processes of these clones. As the desired output, we strive to construct a model that automatically applies refactorings for a large part of the detected duplication problems. This model will be implemented for the Java programming language.
\end{abstract}

@@

Refactoring is used to improve the quality-related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler1999refactoring, wake2004refactoring}, that are integrated into most modern IDE's. However, most of these methods still require a manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research, we will look at to what extend code clones can be automatically refactored.

Code clones can be found anywhere in a codebase. The location of a clone in the code has an impact on how it has to be refactored. Because of this, we will first look at where clones can be found. Using this information we can determine in which locations clones are found the most, on basis of which the refactoring will be prioritized. Apart from that, we will look at different types of clones. A duplicate part in a codebase does not always have to be an exact match with another part to be called a clone. We will look at the definitions of different types of clones and what the opportunities are to refactor duplicate parts of code if it's not an exact match. All of these measurements are performed on a large corpus of open source projects. We lay the main focus on the Java programming language as refactoring opportunities feature paradigm and programming language dependent aspects \cite{choi2011extracting}. However, most practices featured in this research will also be applicable to other object-oriented languages, like C\#.

@@

\section{Clone Detection}
As duplication in source code is a serious problem in many software systems, there are a lot of researches that look into code clones. Many tools have been proposed to detect various types of code clones. Two surveys of modern clone detection tools \cite{sheneamer2016survey, svajlenko2014evaluating} together show an overview of the most-popular clone detection tools up until 2016. To be able to refactor detected clones, it is useful to have the ability to rewrite the AST. We considered a set of clone detection tools for their ability to support the refactoring process automatically. None of the tools we consider seemed completely fit for this purpose, so we decided to implement our own clone detection tool: CloneRefactor\footnote{CloneRefactor (WIP) is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}}.

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. For our tool, we decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio (which is a clone detection tool that's part of TinyPDG: a library for building intraprocedural program dependency graphs for Java programs). We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is useable with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). This is displayed in figure \ref{fig:clonerefactor}.

@@

\subsection{Thresholds}
CloneRefactor works on basis of three thresholds for finding clones:
\begin{enumerate}
  \item \textbf{Number of statements/declarations:} The number of statements/declarations that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of tokens:} The number of tokens (excluding whitespace, end-of-line terminators and comments) that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of lines:} The minimum amount of lines (excluding lines that do not contain any tokens, for tokens same exclusions apply) that should be equal/similar for it to be considered a clone.
\end{enumerate}
Of course, if we put any of these thresholds to zero it won't be taken into account anymore, so not all thresholds have to be used at all times. We consider ``number of lines'' to be the least important, as it highly depends on the programmer of the codebase (and we do not want this kind of dependence!). On basis of manual assessment, we have determined that setting the ``number of statements/declarations'' to 6 ensures that most non-harmful clones are filtered out. On the downside, this also filters out some harmful antipatterns, for instance, if a cloned line has many tokens we might want to consider a clone if it spans less than 6 statements (as a cloned line with many tokens is more harmful than one with few).

@@

\subsubsection{Our measurements}
We have recreated table 3 of Fontana et al \cite{fontana2015duplicated}, but with the following differences in our setup:
\begin{itemize}
  \item We consider clone classes rather than clone pairs.
\item We use different thresholds regarding when a clone should be considered.
\item We seek by statement/declaration rather than SLOC.
\item We test a broader range of projects (they use a set of 50 relatively large projects, we use a large corpus that was assembled by a machine learning algorithm testing java projects on GitHub for quality, which contains projects of all sizes and with differing code quality).
\end{itemize}

Table \ref{table:relations} contains our results regarding the relations between clone instances.

>>>>

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolpceurws}
\usepackage{float}
\usepackage{hyperref}
\usepackage{todonotes}

@@

\begin{abstract}
Duplication in source code can have a \todo{be more specific: what are examples of major impact?}major negative impact on the maintainability of source code. There are several techniques  to merge clones, reduce duplication, improve the design of the code and potentially also reduce the total volume of a software system \todo{so? what is the link between these techniques and what you want to do?}. In this study, we look into the opportunities to aid in the process of refactoring these duplication problems for object-oriented programming languages. We look into \todo{why?} redefinitions for different types of clones that have been used in code duplication research for many years. \todo{turn it around: link the need for refactoring aids to redefining} These redefinitions aim at flagging only those clones useful for refactoring purposes.

Our intermediate results indicate that more than half of the duplication in code is related to each other through inheritance, making it easier to refactor these clones in a clean way. Approximately a fifth of the duplication can be refactored through method extraction, while other clones require other techniques to be applied. Similar future measurements will provide further insight into the contexts where clones occur, and what this means for the refactoring processes. Finally, we strive to construct a \todo{a model cannot apply something} model that automatically applies refactorings for a large part of the detected duplication problems. 
% This model will be implemented for the Java programming language.
\end{abstract}

@@

Refactoring is used to improve the quality-related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler1999refactoring, wake2004refactoring}, that are integrated into most modern IDE's. However, most of these methods still require a manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research, we investigate the extent to which code clones can be automatically refactored.

Code clones can be found anywhere in a codebase. The location of a clone in the code has an impact on how it has to be refactored. Because of this, we will first look at where clones can be found. Using this information we can determine in which locations clones are found the most, on basis of which the refactoring will be prioritized. Apart from that, we will look at different \todo[inline]{make sure to explain that this is a different set of different types, because splitting clones by location in code is also a way to have different types} types of clones. A duplicate part in a codebase does not always have to be an exact match with another part to be called a clone. We will look at the definitions of different types of clones and the opportunities to refactor duplicate parts of code albeit not an exact match. We conduct our research on a large corpus of open source projects. We lay the main focus on the Java programming language as refactoring opportunities feature paradigm and programming language dependent aspects~\cite{choi2011extracting}. However, most practices featured in this research will also be applicable to other object-oriented languages, like C\#.

@@

\section{Clone Detection}
As duplication in source code is a serious problem in many software systems, there are a lot of researches that look into code clones. Many tools have been proposed to detect various types of code clones. Two surveys of modern clone detection tools \cite{sheneamer2016survey, svajlenko2014evaluating} together show an overview of the most-popular clone detection tools up until 2016. To be able to refactor detected clones, it is useful to have the ability to rewrite the AST. We considered a set of clone detection tools for their ability to support the refactoring process automatically. \todo{didn't you have to build a matrix with features etc? where is it? this would be the place for it} None of the tools we consider seemed completely fit for this purpose, so we decided to implement our own clone detection tool: CloneRefactor\footnote{CloneRefactor (WIP) is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}}.

\subsection{CloneRefactor}
\todo{this is not background, this is already your research. Background is what others did and you use as foundation}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. For our tool, we decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio (which is a clone detection tool that's part of TinyPDG: a library for building intraprocedural program dependency graphs for Java programs). We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is useable with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). This is displayed in figure \ref{fig:clonerefactor}.

@@

\subsection{Thresholds}
CloneRefactor works on basis of three thresholds for finding clones:
\begin{enumerate}
  \item \textbf{Number of statements/declarations:} The number of statements/declarations that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of tokens:} The number of tokens (excluding whitespace, end-of-line terminators and comments) that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of lines:} The minimum amount of lines (excluding lines that do not contain any tokens, for tokens same exclusions apply) that should be equal/similar for it to be considered a clone.
\end{enumerate}
Of course, if any threshold is set to zero, it will be ignored, thus not all thresholds have to be used at all times. We consider ``number of lines'' to be the least important, as it highly depends on the programmer of the codebase (and we do not want this kind of dependence!). On basis of manual assessment, we have determined that setting the ``number of statements/declarations'' to 6 ensures that most non-harmful clones are filtered out. On the downside, this also filters out some harmful antipatterns, for instance, if a cloned line has many tokens we might want to consider a clone if it spans less than 6 statements (as a cloned line with many tokens is more harmful than one with few).

@@

\subsubsection{Our measurements}
We use a similar setup to that used in the study of \todo{put here the concise description of the experiment}  (Table~3 in Fontana et al) \cite{fontana2015duplicated}, but with the following differences in our setup:
\begin{itemize}
  \item We consider clone classes rather than clone pairs.
\item We use different thresholds regarding when a clone should be considered.
\item We seek by statement/declaration rather than SLOC.
\item We test a broader range of projects (they use a set of 50 relatively large projects, we use a large corpus that was assembled by a machine learning algorithm testing java projects on GitHub for quality, which contains projects of all sizes and with differing code quality).
\end{itemize}

Table~\ref{table:relations} contains our results regarding the relations between clone instances.
